---
title: "Word Clouds of Indonesian Media Outlet Twitter User Timelines"
author: "Hendy Irawan"
date: "02/04/2015"
output: html_document
---

## R Programming Language Tutorial Videos on YouTube

See [R Programming Language Tutorial Videos by Hendy Irawan](https://www.youtube.com/playlist?list=PLKF12aQHzSaN6i1mz2a1Sg2DIKsXHlVmE).

## Installing Packages: libcurl4-openssl-dev, TwitteR, httpuv, tm, wordcloud, RColorBrewer

Install Ubuntu package `libcurl4-openssl-dev` required by `RCurl` R package:

```
sudo aptitude install libcurl4-openssl-dev
```

Install the R packages:

```{r, eval=FALSE}
install.packages(c('TwitteR', 'httpuv', 'tm', 'wordcloud'))
```

## Setup Twitter OAuth

[Get your Twitter app OAuth consumer credentials](https://apps.twitter.com/), then:

```{r, eval=FALSE}
library(twitter)
setup_twitter_oauth(consumer_key, consumer_secret)
```

## Grab data

```{r, eval=FALSE}
tl_hidcom <- userTimeline('hidcom', n=500)
```

View it as data frame, make sure to convert to `UTF-8` to avoid encoding issues later:

```{r, eval=FALSE}
tl_hidcom.df <- twListToDF(tl_hidcom)
tl_hidcom.df$text <- iconv(tl_hidcom.df$text, to='utf-8')
View(tl_hidcom.df)
```

Get summary:

```{r}
tl_hidcom.df <- read.csv('tl_hidcom_2015-04-02.csv')
summary(tl_hidcom.df)
```

Save the data frame to CSV:

```{r, eval=FALSE}
write.csv(twListToDF(tl_hidcom), 'tl_hidcom.csv')
```

## Make a Corpus

Grab just text column:

```{r}
head(tl_hidcom.df$text)
```

Make a **tm** `Corpus` from the data frame `VectorSource`:

```{r}
library(tm)
tl_hidcom.corpus <- Corpus(VectorSource(tl_hidcom.df$text))
```

Make a `TermDocumentMatrix`, with desired text preprocessors:

```{r}
tl_hidcom.tdm <- TermDocumentMatrix(tl_hidcom.corpus,
  control = list(stripWhitespace = TRUE, tolower = TRUE,
                 stopwords = c('hidayatullah'), removeNumbers = TRUE,
                 removePunctuation = TRUE))
```

Get the matrix from the `TermDocumentMatrix`:

```{r}
tl_hidcom.m <- as.matrix(tl_hidcom.tdm)
tl_hidcom.m[1:10, 1:20]
# View(tl_hidcom.m)
```

Get the word frequencies for `freq > 0`, and sort them (nice way):

```{r}
tl_hidcom.wf <- sort(rowSums(tl_hidcom.m), decreasing=TRUE)
tl_hidcom.wf <- tl_hidcom.wf[tl_hidcom.wf > 0]
tl_hidcom.dm <- data.frame(word=names(tl_hidcom.wf),
                           freq=tl_hidcom.wf)
head(tl_hidcom.dm)
# View(tl_hidcom.dm)
```

or alternatively: (my own convoluted way hehe ;-) )

```{r}
tl_hidcom.dm <- data.frame(word=rownames(tl_hidcom.m),
                           freq=rowSums(tl_hidcom.m))
tl_hidcom.dm <- tl_hidcom.dm[tl_hidcom.dm$freq > 0,]
tl_hidcom.dm <- tl_hidcom.dm[order(tl_hidcom.dm$freq, decreasing=TRUE),]
head(tl_hidcom.dm)
# View(tl_hidcom_dm)
```

## Word Cloud

```{r}
library(wordcloud)
wordcloud(tl_hidcom.dm$word, tl_hidcom.dm$freq,
          random.order=FALSE, colors=brewer.pal(8, 'Dark2'))
```
